---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
#Philemon Ramontal 
#Final Project 
#Research Question: Do females have a higher Literacy Rate than Males around the world? H1:u1>u2, Alternative H0:u1<u2
#Sub Question: 
#Data Description: This dataset represents the 125 countries around the world. It includes the data for overall literacy rate, male literacy rate and female literacy rate. I have added 2 columns (GDP, Unemplement rate) to determine if there is any correlation with the with the literacy rate.

```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
```


#Data Import
```{r}
library(readr)
literacy_rate <- read_csv("literacy_rate.csv")
lit_rate<-literacy_rate
View(lit_rate)
```
#Data Import
```{r}
library(readr)
gdp_improvement_rate <- read_csv("gdp_improvement_rate.csv")
gdp<-gdp_improvement_rate
View(gdp)
```
#Data Import
```{r}
library(readr)
unemployment <- read_csv("unemployment.csv")
View(unemployment)
```
#Data Merge
```{r}
lit_gdp<-merge(lit_rate, gdp, by="country")
View(lit_gdp)
```
#Data merge by "Country"
```{r}
lit_gdp_unemploy<-merge(lit_gdp, unemployment, by="country")
View(lit_gdp_unemploy)
```
#Change Column Names
```{r}
colnames(lit_gdp_unemploy)[colnames(lit_gdp_unemploy)== "growth_rate"] = "GDP"
colnames(lit_gdp_unemploy)[colnames(lit_gdp_unemploy)== "unemployment_percentage"] = "employ_rate"
colnames(lit_gdp_unemploy)[colnames(lit_gdp_unemploy) == "literacy_rate_percent_all"] = "lit_all"
View(lit_gdp_unemploy)
```
#Remove N/A Column
```{r}
lit_gdp_unemploy$X6<-NULL
View(lit_gdp_unemploy)
```

```{r}

```


#StepAIC to determine be variable to use for analysis
```{r}
lit_lm <- lm(lit_all~male_literacy_percent_rate+female_literacy_percent_rate + GDP + employ_rate, data = lit_gdp_unemploy)
step_lit<-step(object = lit_lm, direction = "both")
```
#Anova of StepAIC
```{r}
anova(step_lit)
```
#T.Hypothesis Test question: females have higher literacy rate this males
#u1 > u2 (claim)
#H0: u1=u2
#H1: u1>u2
```{r}
t.test(lit_gdp_unemploy$male_literacy_percent_rate, lit_gdp_unemploy$female_literacy_percent_rate, paired = FALSE, conf.level = 0.95)
```
#P-Value = 0.01422 < alpha 0.05
#Reject Null Hypothesis 
#There is suffieicent evidence to suppport that Females literacy rate is higher than male literacy rate. 
#ggplot of Initally T.test
```{r}
ggplot(data = lit_gdp_unemploy, mapping =  aes(x=lit_gdp_unemploy$male_literacy_percent_rate, y=lit_gdp_unemploy$female_literacy_percent_rate , color = country)) + geom_point() + geom_smooth() + xlab("Male Literacy Rate") + ylab("Female Literacy Rate") + ggtitle("World Literacy Rate")
```
#T.test 2
#Sub Question: Does unemployment have a significant effect on GDP ?
#H0: U1=U2
#H1: U1>U2 (Claim)
```{r}
t.test(lit_gdp_unemploy$GDP, lit_gdp_unemploy$employ_rate, paired = FALSE)
```
#P-Value = 1.941e-10< alpha 0.05
#Reject the null Hypothesis
#There is sufficient evidence to support that unemployment has a significant effect on GDP. 
#ggplot of secondary T.test
```{r}
ggplot(data = lit_gdp_unemploy, mapping = aes(x=lit_gdp_unemploy$GDP, y=lit_gdp_unemploy))+geom_histogram(stat = "identity", position = "dodge") + xlab("GDP") + ylab("Unemployment Rate") + ggtitle("Unemployment vs. GDP")
```
#Regression test
```{r}
lit_lm <- lm(lit_all~male_literacy_percent_rate+female_literacy_percent_rate, lit_gdp_unemploy) 
summary(lit_lm)
```
#P-value = <2e-16
#R-Squared = 0.9932
#Data fits this model very well being that we have a R- squared above .99
#Very Strong Linear relationship
#Regression test 2
```{r}
lit_lm1 <- lm(lit_all~GDP+employ_rate, lit_gdp_unemploy) 
summary(lit_lm1)
```
#P-value = 2e-16
#R-Squaed = 0.1454
#Data does not fit well with model R-Squared is 0.01454 
#Non Linear relationship